An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures

### 2021-04-01 CMPSpaper_writing/CMPSpaper_writing.Rmd
couple of edits

The automated framework  used in this paper was proposed by @aoas. In order to obtain bullet signatures from the x3p files of bullet land engravings, we use the `R` packages `x3ptools` [@x3ptools] and `bulletxtrctr` [@bulletxtrctr]. `x3ptools` \hh{is a package to read, write, and generally, process } x3p files. \hh{The `bulletxtrctr` package implements a pipeline for extracting and comparing signatures from scans of land-engraved areas.}
<!--In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`).--> 
<!--\hh{it's land **engraved** areas, not impressions (that was wrong in Eric's first paper). Impressions are marks from impacts, such as the hit of a hammer or the hit from the breech face. Engravings are scratch marks.}-->
\hh{\autoref{fig:process} gives an overview of all of the steps in the process from scan to signatures.}
\autoref{fig:process}(a) shows a \hh{rendering of a 3d scan of a } bullet land  engraved area. The raised portion of the surface \hh{on the left and right of the scan are parts of the adjacent groove engraved areas (GEAs)}, \hh{the middle area shows a land-engraved area (LEA) with well expressed} striation marks. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land engraving.
<!--\hh{The next section is at an awkward level of detail - there's lots of details, but without the parameter settings it is still not a reproducible description. Could you instead describe things conceptually, and then link to the code that does this particular image?} -->
The white horizontal line in \autoref{fig:process}(a) indicates which cross-section was identified by the algorithm to represent the LEA; \autoref{fig:process}(b) shows the corresponding cross-sectional view. <!--Then we can acquire the crosscut data (`ccdata`) of the scan at this height. --> \hh{XXX passive voice for things that are not part of this paper :)}
Groove engraved areas are removed from the analysis as indicated by the vertical blue lines in \autoref{fig:process}(c) and \autoref{fig:process}(d)). A non-parametric LOESS smooth  [@loess] is fitted to capture the bullet curvature (\autoref{fig:process}(e)) and, finally, the **bullet signature** (\autoref{fig:process}(f)) is obtained as residuals of the cross-section and the fitted smooth. 
\wj{Note that in @cmps bullet signatures are sometimes referred as bullet profiles. However, to avoid confusion, we distinguish the notion of bullet signatures from bullet profiles. Bullet profiles are XXX, while bullet signatures are XXX.}
\hh{Identifiying the groove engraved areas correctly is fundamental for a correct down-stream analysis of the signatures. To allow for a human-in-the middle inspection and intervention we have provided an interactive web-application, implemented as `R` Shiny App [@shiny],}   `bulletinspectR` to identify and correct those errors. 
\hh{An example of the extraction process with corresponding code and parameter settings can be found at XXX.}
Note that \hh{the process of extracting signatures might be different from the one used in @cmps because no code or parameter settings are made available publicly.}

@@

### Conceptual idea of CMPS
\hh{Most algorithms for comparing striation marks are based on the digitized signatures XXX and produce a similarity score (some citations: chumbley score, Ganesh's fix, vorburger's ccf paper, hare).} 
 The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm. The algorithm's main idea  is to take a set of consecutive and non-overlapping basis segments from  the reference profile \wj{might reference profile refer to something more general?}  and  for each segment find the "best" registration position on the comparison profile (the other bullet signature) with respect to their cross-correlation values. \hh{From a comparison of these registration positions, a } **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score. 
\hh{High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair. Low scores between pairs of signatures are attributed to different source pairs.} 
\hh{XXX What threshold did @cmps identify as the boundary between different source and same source? And what was the error rate in their example?}
\wj{they didn't talk about a specific threshold or error rates}
 The CMPS algorithm can assist firearm examiners with drawing a conclusion \hh{about the source of a comparison pair}. \hh{Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the } CMPS algorithm. 
Thus, in this paper we present an open-source implementation of the CMPS algorithm \hh{in the R package CMPS available from XXX github / CRAN}. 
With our implementation, one can compute the CMPS score of a comparison using the following code:

>>>>

The automated framework  used in this paper was proposed by @aoas. In order to obtain bullet signatures from the x3p files of bullet land engravings, we use the `R` packages `x3ptools` [@x3ptools] and `bulletxtrctr` [@bulletxtrctr]. `x3ptools` \hh{is a package to read, write, and generally, process } x3p files. \hh{The `bulletxtrctr` package implements a pipeline for extracting and comparing signatures from scans of land-engraved areas.}
<!--In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`).--> 
<!--\hh{it's land **engraved** areas, not impressions (that was wrong in Eric's first paper). Impressions are marks from impacts, such as the hit of a hammer or the hit from the breech face. Engravings are scratch marks.}-->
\hh{\autoref{fig:process} gives an overview of all of the steps in the process from scan to signatures.}
\autoref{fig:process}(a) shows a \hh{rendering of a 3d scan of a } bullet land  engraved area. The raised portion of the surface \hh{on the left and right of the scan are parts of the adjacent groove engraved areas (GEAs)}, \hh{the middle area shows a land-engraved area (LEA) with well expressed} striation marks. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land engraving.
<!--\hh{The next section is at an awkward level of detail - there's lots of details, but without the parameter settings it is still not a reproducible description. Could you instead describe things conceptually, and then link to the code that does this particular image?} -->
The white horizontal line in \autoref{fig:process}(a) indicates which cross-section was identified by the algorithm to represent the LEA; \autoref{fig:process}(b) shows the corresponding cross-sectional view. <!--Then we can acquire the crosscut data (`ccdata`) of the scan at this height. --> \hh{XXX passive voice for things that are not part of this paper :)}
Groove engraved areas are removed from the analysis as indicated by the vertical blue lines in \autoref{fig:process}(c) and \autoref{fig:process}(d)). A non-parametric LOESS smooth  [@loess] is fitted to capture the bullet curvature (\autoref{fig:process}(e)) and, finally, the **bullet signature** (\autoref{fig:process}(f)) is obtained as residuals of the cross-section and the fitted smooth. 
Note, that in @cmps bullet signatures are  referred to as bullet profiles. However, to avoid confusion, we distinguish the notion of bullet signatures from bullet profiles. \wj{Bullet profiles are \hh{shown in panels (b), (c), and (d) of \autoref{fig:process}}, while \autoref{fig:process}(f) shows the corresponding bullet signature.} 
\hh{Identifiying the groove engraved areas correctly is fundamental for a correct down-stream analysis of the signatures. To allow for a human-in-the middle inspection and intervention we have provided an interactive web-application, implemented as `R` Shiny App [@shiny],}   `bulletinspectR` to identify and correct those errors. 
\hh{An example of the extraction process with corresponding code and parameter settings can be found at XXX.}
Note that \hh{the process of extracting signatures might be different from the one used in @cmps because no code or parameter settings are made available publicly.}

@@

### Conceptual idea of CMPS
\hh{Most algorithms for comparing striation marks are based on the digitized signatures XXX and produce a similarity score (some citations: chumbley score, Ganesh's fix, vorburger's ccf paper, hare).} 
 The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm. The algorithm's main idea  is to take a set of consecutive and non-overlapping basis segments from  the reference   and  for each segment find the "best" registration position on the comparison  (the other bullet signature) with respect to their cross-correlation values. \hh{From a comparison of these registration positions, a } **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score. 
\hh{High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair. Low scores between pairs of signatures are attributed to different source pairs.} 
\hh{XXX What threshold did @cmps identify as the boundary between different source and same source? And what was the error rate in their example?}
\wj{they didn't talk about a specific threshold or error rates} \hh{XXX Is there a way to infer the error based on the example? ... 'while the authors did not mention a specific threshold of the CMPS to distinguish between same-source and different-source comparison, a CMPS value of Y results in an error rate of Z in the example.'}
 The CMPS algorithm can assist firearm examiners with drawing a conclusion \hh{about the source of a comparison pair}. \hh{Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the } CMPS algorithm. 
Thus, in this paper we present an open-source implementation of the CMPS algorithm \hh{in the R package CMPS available from XXX github / CRAN}. 
With our implementation, one can compute the CMPS score of a comparison using the following code:
